---
title: "Transit Costs Project"
description: |
  A short description of the post.
author:
  - name: Ethan Tenison
    url: https://twitter.com/ethantenison
date: 01-04-2022
output:
  distill::distill_article:
    self_contained: false
---

# Transit Costs Project 

This data comes from the [Transit Costs Project](https://transitcosts.com/) via `#tidytuesday` on Twitter.

>Why do transit-infrastructure projects in New York cost 20 times more on a per kilometer basis than in Seoul? We investigate this question across hundreds of transit projects from around the world. We have created a database that spans more than 50 countries and totals more than 11,000 km of urban rail built since the late 1990s. We will also examine this question in greater detail by carrying out six in-depth case studies that take a closer look at unique considerations and variables that aren’t easily quantified, like project management, governance, and site conditions.

>The goal of this work is to figure out how to deliver more high-capacity transit projects for a fraction of the cost in countries like the United States. Additionally, we hope that our site will be a useful resource for elected officials, planners, researchers, journalists, advocates, and others interested in contextualizing transit-infrastructure costs and fighting for better projects.

One of the reason why I chose it was because the city I live in, Austin, approved a massive expansion of its rail transit-system. From the beginning of the planning process (a different version was defeat once before), it was highly controversial partly because of the cost. Given the congestion on Austin's roads already, it doesn't seem like the city has any other viable option to alleviate the problem. 


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)


```

```{r data, echo = TRUE}

library(tidytuesdayR)
library(tidyverse
        )

tuesdata <- tidytuesdayR::tt_load(2021, week = 2)

transit_cost <- tuesdata$transit_cost
glimpse(transit_cost)

```


# Average cost per country

```{r total_cost}

total <- transit_cost |> 
  group_by(country) |> 
  summarize(average_cost = mean(cost_km_millions)) |> 
  filter(!is.na(average_cost),
         !is.na(country)) |> 
  mutate(us = case_when(
    country == "US" ~ "1",
    country == "CN" ~ "1",
    TRUE ~ "0"
  ))

theme_set(theme_classic())

ggplot(total, aes(x = reorder(country, -average_cost), y = average_cost, fill = us)) +
  geom_bar(stat = "identity") +
  scale_fill_manual( values = c( "1"="red", "0"="darkgray" ), guide = FALSE )+
  labs( x = "Country",
        y = "Average cost per km",
        title = "Average rail cost per km by country",
        subtitle = "(Millions $USD)", 
        caption = "Source: Transit Costs Project") +
  geom_curve(x = 8.5, y = 1100,
             xend = 2, yend = 1000,
             color = "black",
             curvature = 0.2,
             size = 1.5,
             arrow = arrow()) +
  annotate(
      "text",
      x = 18,
      y = 1000,
      label = expression("The United States pays the most\n in the world per kilometer of rail\n at $1211.46 million per kilometer"),
      size = 3.5
   ) +
  geom_curve(x = 40, y = 350,
             xend = 38, yend = 200,
             color = "black",
             curvature = 0.2,
             size = 1.5,
             arrow = arrow()) +
  annotate(
      "text",
      x = 48,
      y = 350,
      label = expression("China only spends on\n average $184.39 million"),
      size = 3.5
   ) +
  annotate("text", 
           x = 1.2,
           y = -30,
           label = "US", 
           size = 3.5) +
  annotate("text", 
           x = 38,
           y = -30,
           label = "CN", 
           size = 3.5) +
  theme(
    axis.text.x = element_blank(),
    title = element_text(hjust = 0.5),
    axis.ticks.x=element_blank(),
    axis.line.x = element_blank(),
    axis.line.y = element_blank()
  )

```

# Total Lines per country 

```{r total_lines}

total_lines <- transit_cost |> 
  group_by(country) |>
  count() |> 
  filter(!is.na(country)) |> 
  mutate(us = case_when(
    country == "CN" ~ "1",
    country == "US" ~ "1",
    TRUE ~ "0"
  ))

ggplot(total_lines, aes(x = reorder(country, -n), y = n, fill = us)) +
  geom_bar(stat = "identity") +
  scale_fill_manual( values = c( "1"="red", "0"="darkgray" ), guide = FALSE ) +
  labs( x = "Countries",
        y = "Rail lines built",
        title = "Total rail lines built per country",
        subtitle = "(1990 - )",
        caption = "Source: Transit Costs Project") +
  theme(
    axis.text.x = element_blank(),
    title = element_text(hjust = 0.5),
    axis.ticks.x=element_blank() ,
    axis.line.x = element_blank(),
    axis.line.y = element_blank()
  ) +
  geom_curve(x = 10, y = 175,
             xend = 2, yend = 150,
             color = "black",
             curvature = 0.2,
             size = 1.5,
             arrow = arrow()) +
  annotate(
      "text",
      x = 17,
      y = 170,
      label = expression("By far, China built the\n most rail since 1990."),
      size = 3.5
   ) +
  annotate("text", 
           x = 1.3,
           y = -8,
           label = "CN", 
           size = 3.5) +
  annotate("text", 
           x = 8,
           y = -8,
           label = "US", 
           size = 3.5)

```



